---
title: "p8105_hw2_ph2538"
author: "Pei Yang Hsieh"
date: 2018-09-28
output: github_document
---

```{r setup}
library(tidyverse)
library(ggplot2)
library(readxl)
library(devtools)

```

The purpose of this file is to complete homework 2.

## Problem 1

### Loading and Cleaning Data

```{r load_clean_1}
#Load and clean data: retain line, station, name, station latitude/longitude, routes served, entry, vending, entrance type, and ADA compliance
#Convert entry variable from character (YES vs NO) to logical variable

NYC_transit_data = read_csv(file = "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>%
  janitor::clean_names() %>%
  select(line, station_name, station_latitude, station_longitude, starts_with("route"), entry, vending, entrance_type, ada) %>%
  mutate(entry = recode(entry, 'YES' = "TRUE", 'NO' = "FALSE"))
```

Write a short paragraph about this dataset - explain briefly what variables the dataset contains, describe your data cleaning steps so far, and give the dimension (rows x columns) of the resulting dataset. Are these data tidy?

This dataset includes variables on line, station name, station latitude and longitude, routes, entry, vending, entrance type, and ADA compliance. My data cleaning steps have included loading the data and cleaning the variable names, so they are all in snake case. Next, I selected only the variables that I needed, discarding the rest. Lastly, I converted the entry variable from a character to a logical variable.

The dimensions of the resulting dataset are `r dim(NYC_transit_data)`

These data are not tidy because the the routes are spread across multiple columns, with many missing values.

## Problem 2

```{r load_2}
trash_wheel_data = read_excel("./data/HealthyHarborWaterWheelTotals2017-9-26.xlsx")
```


## Problem 3

```{r load_3}
#devtools::install_github("p8105/p8105.datasets")
library(p8105.datasets)

data(brfss_smart2010)

brfss_smart2010

#df_name = brfss_smart2010
```

